Advanced Normalization Tools (ANTS)

### 2011-03-20 Documentation/Tex/ants.tex
ENH: documentation on metric weights was unclear.  this should help some.
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@813 fd2da0a4-d968-48df-9cb5-d8c496287cbf

\usepackage[dvips]{graphicx}
%\usepackage{listing}						
\usepackage{listings}	
\usepackage{wrapfig}
\usepackage{amssymb,amsmath}
\usepackage{multirow,booktabs,ctable,array}

@@

\subsubsection{Deformation Field File}
ANTS writes the deformation field using one file for the deformation in each dimension. 
For example, if the deformation field is named ABCD.nii.gz and exists in three dimensions, the corresponding files are:
ABCDxvec.nii.gz, 
ABCDyvec.nii.gz, 
ABCDzvec.nii.gz (for 3D).  Note that ANTS programs {\em assume} the $*$vec part of the name will exist.  Thus, 
one refers to the warp file as  ABCD.nii.gz , dropping the $*$vec from the name. 
The displacement for each voxel is defined in physical image space.  If you want to 
see the details of how a warp is achieved, look at the code in itkWarpImageMultiTransform.txx. 
One issue to be aware of is that if you are directly indexing a vector component image in 
voxel coordinates, then you need to apply the image's direction matrix to the vector.  
An example of this is in ANTSJacobian.cxx, the TransformVector function.

\subsubsection{Labeled Point Sets}
There are three formats supported
\begin{itemize}
\item  VTK format: point + scalar label in a .vtk file. {\bf FIXME NEED MORE DETAIL -- what type of vtk data ?  polydata?}.
\item  Labeled point sets are saved as an image volume with the intensity = 1,2,3 for label = 1,2,3. For example, 
figure~\ref{fig:frown} shows 2D images representing a 3-class labeled point set. 
\item Avants Format. In this simple format, the point sets are saved as a text file. The first and last line is always 0 0 0 0. Each line in between for a point as X Y Z L (L is the label.)
\end{itemize}
The names and formats of the output files are both specified by -o option to the ANTS executable. 
For example, for in 3D with -o RegA2B.nii.gz , the typical output files are:
Affine transform file: 
RegA2BAffine.txt 
Warping deformation field: 
RegA2BWarpxvec.nii.gz, RegA2BWarpyvec.nii.gz, RegA2BWarpzvec.nii.gz
Inverse warping deformation field:
RegA2BInverseWarpxvec.nii.gz, RegA2BInverseWarpyvec.nii.gz, RegA2BInverseWarpzvec.nii.gz.  
To apply the affine transform and the deformation field to the moving image, use WarpImageMultiTransform. 
For example: to apply the transform to the moving image:

@@

\subsection{The ImageMath Executable} This is a multi-purpose program
that has the following syntax:  ``ImageMath ImageDimension
outputfilename   Operation  InputFileName  parameters ``.  Most basic
scalar image operations -- and some tensor operations -- may be
performed with this program.  Some operations output text files, some
output images and some output only to the terminal.   ImageMath allows
one to multiply images together (m), to negate images (Neg), to take
an image to a power (pow), to test the invertibility of
transformations (InvId), to compute the fractional anisotropy of an
image (TensorFA) and to perform Markov Random Field segmentation with
or without priors (Segment).   Many other operations are available.
Like all other ANTS programs, one may call ImageMath from the command
line to see all of its options.   ImageMath is used heavily in ANTS
scripts.   You can call it, in compination with grep or less, to attempt to
find some of its capabilities.

\subsection{ANTS Scripts} 
The ANTS/Scripts directory contains (hopefully) user-friendly
wrappings of ANTS tools that enable higher-level error checking and
combinations of basic ANTS functions.  These scripts are called as
\verb sh ~ \verb antsscriptname.sh ~ and provide usage when called
from the command line.  For instance, try \verb sh ~ \verb ants.sh .
Additionally, if you run ants in tcsh shells, you will need to call it
as follows:
\begin{verbatim}
~/Code/ANTSSourceForgeSVN/bin/ANTS 2 -m CC\[r16slice.nii,r64slice.nii,1,2\]
\end{verbatim}
instead of
\begin{verbatim}
~/Code/ANTSSourceForgeSVN/bin/ANTS 2 -m CC[r16slice.nii,r64slice.nii,1,2]
\end{verbatim}
because tcsh does no interpret the brackets as intended.  Using the
bash shell script interface also avoids this.  The key change involves using the slash to force tcsh to 
interpet the bracket literally. 
\newpage
\section{ImageRegistration with ANTS}
\begin{wrapfigure}{r}{0.5\textwidth}
\framebox[0.5\textwidth]{
\begin{minipage}[t]{0.45\textwidth}
{\bf Quick Start:} {\color{blue}call ants.sh (from ANTS/Scripts/) to get usage and apply a normalization to some of your existing data.  It is instructional to read the script, modify some of the parameters and re-run to witness the effect of your changes -- image registration is an art as well as science.  Many other ``ready to go'' scripts are available in ANTS/Scripts. The user must set his/her ANTSPATH environment variable  -- which points to the location of ANTS binaries -- within these scripts or in their environment.  Note: All ANTS programs provide usage when called from the command line.  Most require the image dimension to be specified as the first parameter.  E.g. 
  ImageMath ImageDimension 
where ImageDimenion is 2 or 3.}
\end{minipage}}
\end{wrapfigure}
There are two general applications of image registration. The first application is transforming labeled data from a template image space into an individual space. This strategy is important when appearance alone is not enough to locate a structure as, for example, in the case of hippocampus segmentation. The template provides a prediction of the hippocampus-amygdala boundary.
The second application operates in the "inverse" direction of the first: instead of mapping template to individual, we map individual(s) to the template. Voxel-based population studies of either functional or structural variables depend on mapping to a template space. The common coordinate system enables a statistical evaluation of the likelihood of consistent activation across a group or, in other contexts, the differences in anatomy between two groups.

@@

Thus, ANTS enables many operating points from both the transformation and appearance domains such that users may make choices appropriate for their problems. The ANTS command-line syntax, shown in figure~\ref{fig:cmd}, reflects these operating points and the various components that interact in the optimization.
\begin{figure}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax.pdf} 
\itkcaption[ANTS Command Line]{The relationship between the variational 
form that defines the optimization goals and the ANTS command line.  The 
SyN option implements the method in \cite{Avants2008b} and evaluated in 
\cite{Klein2009}.}
\vspace{-0.1in}
\label{fig:cmd}
\end{figure}
ANTS may be used to navigate the transformation and similarity metric space. 
\subsection{Use Your Header!} ANTS uses the direction/orientation, spacing
and origin in its definitions of the mapping between images.  That is,
ANTS uses the nifti standard defintions of an image space.  So, the
image orientation (direction), origin and spacing are important!  You
may ``check'' for consistency between two image header definitions by
using: ImageMath 3 Image2repaired.nii.gz CompareHeadersAndImages
Image1.nii.gz Image2.nii.gz \newline This will compare Image2 to
Image1 and ``fix'' the header of Image2, writing out to
Image2repaired.

@@

\noindent{\bf Diffeomorphic Transformations.}  The elastic 
mapping space -- as indicated above -- may prove inadequate 
for some large deformation mapping scenarios.  Figures~\ref{fig:antshier} 
illustrate the changing performance one may get in switching from affine 
to elastic to diffeomorphic normalization.  
\begin{figure}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSsyn.pdf} 
\itkcaption[ANTS Symmetric Normalization]{This example shows the benefit of the symmetric normalization model -- invertibility, symmetry, highly deformable and accurate registration.  The transformation model used here is highlighted in red.}
\vspace{-0.1in}
\label{fig:antssyn}
\end{figure}
Figure~\ref{fig:antssyn} shows 
how one may use ANTS to achieve a state-of-the-art diffeomorphic mapping. 
The ANTS diffeomorphic model chosen for this example -- symmetric normalization \cite{Avants2008b,Kim2008} -- 
is invariant to the order of the input images (although the affine mapping is not).  
An additional advantage of the diffeomorphic model over the elastic model is 
that both forward and inverse transformations are computed thus allowing one 
to deform fixed to moving and also moving to fixed.   The transformation model 
chosen here -- \verb SyN[0.25] -- may be replaced with other 
diffeomorphic transformation models.  The most general is global-in-time SyN via \verb SyN[0.25,2,0.01]   , 
where the time step for integration is 0.01 (lower is more accurate).  A fast approximate 
solution may be gained through 
exponential mapping via \verb Exp[0.25,10] , where 10 integrations points are used. 
Update schemes for diffeomorphic optimization are similar to the elastic case 
and are described elsewhere \cite{Avants2008b,ANTStheory}.

\subsection{ANTS Similarity Terms}
Here is a script that will let you experiment with different similarity term 
and transformation model combinations.  A few are listed, with parameters 
that are a reasonable starting point.
\begin{verbatim}
II=r16slice.nii # change to your own images 
JJ=r64slice.nii 
DIM=2 # two-dimensional images 
ITS=" -i 100x100x30 " # 3 optimization levels 
TSYNWITHTIME=" -t SyN[0.5,5,0.01]  -r Gauss[3,0] "
TGREEDYSYN=" -t SyN[0.25]  -r Gauss[3,0] "
TELAST=" -t Elast[1] -r Gauss[0.5,3] " 
TEXP="  -t Exp[0.5,10] -r Gauss[0.5,3] " 
LABELGUIDED=" -m PSE[${II},${JJ},${II},${JJ},0.75,0.1,25,0,10] "
INTMSQ=" -m MSQ[${II},${JJ},1,0] "
INTMI=" -m MI[${II},${JJ},1,32] "
INTCC=" -m CC[${II},${JJ},1,4] "
NAME=TEST
INT=$INTPR
TRAN=$TSYN
ANTS $DIM -o $NAME $ITS  $TRAN  $INT
INVW=" -i ${NAME}Affine.txt ${NAME}InverseWarp.nii.gz "
FWDW=" ${NAME}Warp.nii.gz ${NAME}Affine.txt "
WarpImageMultiTransform  2 ${II} a.nii.gz -R ${JJ} $INVW
WarpImageMultiTransform  2 ${JJ} b.nii.gz -R ${II} $FWDW
\end{verbatim}
Convergence occurs under two conditions: (1) either the maximum number of 
iterations are reached at a given optimization level or (2) the slope 
of the change in the optimization objective is negative or very small. 
{\em Note: ANTS does not like spaces between brackets in the call to the metrics.}

\subsection{Choosing a Metric}
\input{metric.tex}

@@

\item Copy the file ANTS/Scripts/buildtemplate.sh into that directory.  

@@

\verb  bash ~ \verb buildtemplateparallel.sh -h ~ to get usage. 

@@

The script offers an option ( -c 1  for qsub , -c 2 for voxbo )  so that you can send the whole thing to distributed computing. 

@@

This succeeded for all subjects but one.  This subject had severe neurodegeneration which caused "under-normalization" to result.  A larger deformation mapping was required and so we increased the maximum number of iterations (the "-i" parameter).  One might also increase the gradient-descent step-size (SyN[0.25] => SyN[0.5]).   Large deformation mapping is challenging not only because of the amount of deformation, but also because the "true" solution is difficult to find.  Keeping this in mind, we also increased the "span" (radius) of the correlation window in the similarity metric (which also increases computation time). These modified parameters succeeded: 

@@

\verb Atropos tool, which is still under development.  We apply 

@@

@@

@@

\itkcaption[ANTS Segmentation]{ANTS extends basic Markov Random Field
  regularized Gaussian model tissue segmentation to include priors,

@@

  segmentation.  The ANTS program N3BiasFieldCorrection is less

@@

\item Preprocess images -- bias correction (e.g. via the \verb
  N4BiasFieldCorrection program), segmentation (via \verb Atropos ), etc.  Potentially construct an optimal template. 

>>>>

% \usepackage[dvips]{graphicx}
%\usepackage{listing}						
\usepackage{listings}	
\usepackage{wrapfig}
\usepackage{amssymb,amsmath}
\usepackage{multirow,booktabs,ctable,array}

@@

\subsubsection{Deformation Field File}
ANTS writes the deformation field into a multi-component image (latest
release).  Thus, 
one refers to the warp file as ABCD.nii.gz which, in 3D, contains 3 components.  If you want to 
see the details of how a warp is achieved, look at the code in itkWarpImageMultiTransform.txx. 
One issue to be aware of is that if you are directly indexing a vector component image in 
voxel coordinates, then you need to apply the image's direction matrix to the vector.  
An example of this is in ANTSJacobian.cxx, the TransformVector function.

\subsubsection{Labeled Point Sets}
There are three formats supported
\begin{itemize}
\item  VTK format: point + scalar label in a .vtk file. {\bf FIXME NEED MORE DETAIL -- what type of vtk data ?  polydata?}.
\item  Labeled point sets are saved as an image volume with the intensity = 1,2,3 for label = 1,2,3. For example, 
figure~\ref{fig:frown} shows 2D images representing a 3-class labeled point set. 
\item Avants Format. In this simple format, the point sets are saved as a text file. The first and last line is always 0 0 0 0. Each line in between for a point as X Y Z L (L is the label.)
\end{itemize}
The names and formats of the output files are both specified by -o option to the ANTS executable. 
For example, for in 3D with -o RegA2B.nii.gz , the typical output files are:
Affine transform file: 
RegA2BAffine.txt 
Warping deformation field: 
RegA2BWarp.nii.gz.
Inverse warping deformation field:
RegA2BInverseWarp.nii.gz.  
To apply the affine transform and the deformation field to the moving image, use WarpImageMultiTransform. 
For example: to apply the transform to the moving image:

@@

\subsection{The ImageMath Executable} This is a multi-purpose program
that has the following syntax:  ``ImageMath ImageDimension
outputfilename   Operation  InputFileName  parameters ``.  Most basic
scalar image operations -- and some tensor operations -- may be
performed with this program.  Some operations output text files, some
output images and some output only to the terminal.   ImageMath allows
one to multiply images together (m), to negate images (Neg), to take
an image to a power (pow), to test the invertibility of
transformations (InvId), to compute the fractional anisotropy of an
image (TensorFA) and to perform many other feats of image processing strength.   
Like all other ANTS programs, one may call ImageMath from the command
line to see all of its options.   ImageMath is used heavily in ANTS
scripts.   You can call it, in compination with grep or less, to attempt to
find some of its capabilities.

\subsection{ANTS Scripts} 
The ANTS/Scripts directory contains (hopefully) user-friendly
wrappings of ANTS tools that enable higher-level error checking and
combinations of basic ANTS functions.  These scripts are called as
\verb bash ~ \verb antsscriptname.sh ~ and provide usage when called
from the command line.  For instance, try \verb bash ~ \verb ants.sh .
Additionally, if you run ants in tcsh shells, you will need to call it
as follows:
\begin{verbatim}
~/Code/ANTSSourceForgeSVN/bin/ANTS 2 -m CC\[r16slice.nii,r64slice.nii,1,2\]
\end{verbatim}
instead of
\begin{verbatim}
~/Code/ANTSSourceForgeSVN/bin/ANTS 2 -m CC[r16slice.nii,r64slice.nii,1,2]
\end{verbatim}
because tcsh does not interpret the brackets as intended.  Using the
bash shell script interface also avoids this.  The key change involves using the slash to force tcsh to interpet the bracket literally. 
\newpage
\section{ImageRegistration with ANTS}
\begin{wrapfigure}{r}{0.5\textwidth}
\framebox[0.5\textwidth]{
\begin{minipage}[t]{0.45\textwidth}
{\bf Quick Start:} {\color{blue}call ants.sh (from ANTS/Scripts/) to get usage and apply a normalization to some of your existing data.  It is instructional to read the script, modify some of the parameters and re-run to witness the effect of your changes -- image registration is an art as well as science.  Many other ``ready to go'' scripts are available in ANTS/Scripts. The user must set his/her ANTSPATH environment variable  -- which points to the location of ANTS binaries -- within these scripts or in their environment.  Note: All ANTS programs provide usage when called from the command line.  Most require the image dimension to be specified as the first parameter.  E.g. 
  ImageMath ImageDimension 
where ImageDimenion is 2, 3 or 4.}
\end{minipage}}
\end{wrapfigure}
There are two general applications of image registration. The first application is transforming labeled data from a template image space into an individual space. This strategy is important when appearance alone is not enough to locate a structure as, for example, in the case of hippocampus segmentation. The template provides a prediction of the hippocampus-amygdala boundary.
The second application operates in the "inverse" direction of the first: instead of mapping template to individual, we map individual(s) to the template. Voxel-based population studies of either functional or structural variables depend on mapping to a template space. The common coordinate system enables a statistical evaluation of the likelihood of consistent activation across a group or, in other contexts, the differences in anatomy between two groups.

@@

Thus, ANTS enables many operating points from both the transformation and appearance domains such that users may make choices appropriate for their problems. The ANTS command-line syntax, shown in figure~\ref{fig:cmd}, reflects these operating points and the various components that interact in the optimization.
\begin{figure}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax.pdf} 
\itkcaption[ANTS Command Line]{The relationship between the variational 
form that defines the optimization goals and the ANTS command line.  The 
SyN option implements the method in \cite{Avants2008b} and evaluated in 
\cite{Klein2009}.}
\vspace{-0.1in}
\label{fig:cmd}
\end{figure}
ANTS may be used to navigate the transformation and similarity metric space. 
\subsection{Use Your Header!} ANTS uses the direction/orientation, spacing
and origin in its definitions of the mapping between images.  That is,
ANTS uses the nifti standard defintions of an image space.  So, the
image orientation (direction), origin and spacing are important!  The
program 'PrintHeader' will give you information about your image's header.

@@

\noindent{\bf Diffeomorphic Transformations.}  The elastic 
mapping space -- as indicated above -- may prove inadequate 
for some large deformation mapping scenarios.  Figures~\ref{fig:antshier} 
illustrate the changing performance one may get in switching from affine 
to elastic to diffeomorphic normalization.  
\begin{figure}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSsyn.pdf} 
\itkcaption[ANTS Symmetric Normalization]{This example shows the benefit of the symmetric normalization model -- invertibility, symmetry, highly deformable and accurate registration.  The transformation model used here is highlighted in red.}
\vspace{-0.1in}
\label{fig:antssyn}
\end{figure}
Figure~\ref{fig:antssyn} shows 
how one may use ANTS to achieve a state-of-the-art diffeomorphic mapping. 
The ANTS diffeomorphic model chosen for this example -- symmetric normalization \cite{Avants2008b,Kim2008} -- 
is invariant to the order of the input images (although the affine mapping is not).  
An additional advantage of the diffeomorphic model over the elastic model is 
that both forward and inverse transformations are computed thus allowing one 
to deform fixed to moving and also moving to fixed.   The transformation model 
chosen here -- \verb SyN[0.25] -- may be replaced with other 
diffeomorphic transformation models.  The step-size (0.25) impacts
accuracy.  Smaller is generally more accurate but takes more time to
compute and may not capture as much deformation if the optimization is
caught in a local minimum.  The most general is global-in-time SyN via \verb SyN[0.25,2,0.01]   , 
where the time step for integration is 0.01 (lower is more accurate).  A fast approximate 
solution may be gained through 
exponential mapping via \verb Exp[0.25,10] , where 10 integrations points are used. 
Update schemes for diffeomorphic optimization are similar to the elastic case 
and are described elsewhere \cite{Avants2008b,ANTStheory}.

\subsection{ANTS Similarity Terms and Other Functions}
Here is a script that will let you experiment with different similarity term 
and transformation model combinations.  A few are listed, with parameters 
that are a reasonable starting point.
\begin{verbatim}
# The user should define two images.  here we use ANTS reference data
II=r16slice.nii.gz 
JJ=r64slice.nii.gz
# 
OUTPUTNAME=OUT  # define the output prefix 
ITS=" -i 100x100x30 " # 3 optimization levels

# different transformation models you can choose
TSYNWITHTIME=" -t SyN[0.25,5,0.01] -r Gauss[3,0] " # spatiotemporal (full) diffeomorphism
TGREEDYSYN=" -t SyN[0.15] -r Gauss[3,0] "          # fast symmetric normalization
TELAST=" -t Elast[1] -r Gauss[0.5,3] "             # elastic 
TEXP=" -t Exp[0.5,10] -r Gauss[0.5,3] "            # exponential

# different metric choices for the user 
INTMSQ=" -m MSQ[${II},${JJ},1,0] "
INTMI=" -m MI[${II},${JJ},1,32] " 
INTCC=" -m CC[${II},${JJ},1,4] " 
#
#
# these are the forward and backward warps. 
#
INVW=" -i ${OUTPUTNAME}Affine.txt ${OUTPUTNAME}InverseWarp.nii.gz " 
FWDW=" ${OUTPUTNAME}Warp.nii.gz ${OUTPUTNAME}Affine.txt " 
#
INT=$INTCC  # choose a  metric ( here , cross-correlation )
TRAN=$TGREEDYSYN  # choose a transformation 
if [[ $OPT == "Reg" ]] ; then

# run the registration
${ANTSPATH}/ANTS $DIM -o $OUTPUTNAME $ITS $TRAN $INT 
# this is how you apply the output transformation 
${ANTSPATH}/WarpImageMultiTransform $DIM ${II} ${OUTPUTNAME}IItoJJ.nii.gz -R ${JJ} $INVW 
${ANTSPATH}/WarpImageMultiTransform $DIM ${JJ} ${OUTPUTNAME}JJtoII.nii.gz -R ${II} $FWDW

elif [[ $OPT == "RegSeg" ]] ; then

# try something different -- segment the images and use them in ANTS 
# to run multivariate registration.
IISEG=${OUTPUTNAME}A_seg.nii.gz
JJSEG=${OUTPUTNAME}B_seg.nii.gz
#
# get masks for atropos 
#
${ANTSPATH}ThresholdImage $DIM $II $IISEG 1 1.e9
${ANTSPATH}ThresholdImage $DIM $JJ $JJSEG 1 1.e9
AtroposParams=" -d $DIM -m [0.1,1x1] -c [5,0] -i kmeans[3] "
${ANTSPATH}/Atropos $AtroposParams -a $II -o $IISEG  -x $IISEG 
${ANTSPATH}/Atropos $AtroposParams -a $JJ -o $JJSEG  -x $JJSEG 
# compute some segmentations and use them in labelguided mapping 
LABELGUIDED=" -m PSE[${II},${JJ},${IISEG},${JJSEG},0.75,0.1,25,0,10] "  
#
${ANTSPATH}/ANTS $DIM -o ${OUTPUTNAME} $ITS $TRAN $INT $LABELGUIDED 
${ANTSPATH}/WarpImageMultiTransform $DIM ${II} ${OUTPUTNAME}IItoJJ.nii.gz -R ${JJ} $INVW 
${ANTSPATH}/WarpImageMultiTransform $DIM ${JJ} ${OUTPUTNAME}JJtoII.nii.gz -R ${II} $FWDW

elif [[ $OPT == "BTP" ]] ; then 
#
# now build a template from the two input images .
# in a real application, you would modify the call to btp 
# such that it's more appropriate for your data.  
# call buildtemplateparallel.sh -h to get usage.
# 
${ANTSPATH}buildtemplateparallel.sh -d 2  -o ${OUTPUTNAME}BTP -c 0 $II $JJ
fi 
\end{verbatim}
Convergence occurs under two conditions: (1) either the maximum number of 
iterations are reached at a given optimization level or (2) the slope 
of the change in the optimization objective is negative or very small.

\subsection{Choosing a Metric}
ANTS supports both volumetric registration and point set registration. The image / point set similarity metrics in ANTS are unified in the form of a function on the images or the point sets:       $$\textbf{Similarity}[\textbf{fixedImage}, \textbf{movingImage}, \textbf{weight}, \textbf{parameters}].$$

The similarity type for the deformation transformation is specified by
the \verb -m  option, which contains two parts: simarity type and the parameters inside the brackets. (Note: no white spaces exist between parameters.) The possible similarity metrics for volumetric images are: 
\begin{itemize}
    \item Cross correlation estimate:  \verb " -m  CC[fixedImage,movingImage,weight,radius]". This metric works
      well for intra-modality and some inter-modality image registration. For example, \verb"-m CC[fixed.nii,moving.nii,1,5]" specifies:
    \begin{itemize}
        \item the fixed image: fixed.nii
        \item the moving image: moving.nii             
        \item weight for this metric is 1 (the weights are relative to
          the weights on the $N$ other metrics passed to ANTS --- $N$
          is unlimited).  So, the weight, $w_i$ on the $i^{th}$ metric
          will be  $w_i=w_i/ ( \sum_i  w_i )$.
        \item the region radius for computing cross correlation is 5
    \end{itemize}

    \item Mutual information: \verb"-m MI[fixedImage,movingImage,weight,number-of-histogram-bins]". This metric works both well for intra-modality and inter-modality image registration. For example, the first three parameters in \verb"-m MI[fixed.nii,moving.nii,1,32]" is similar to the example above in cross correlation, except that the last parameter means that the number of bins in computing mutual information is 32.

    \item PR: \verb"-m PR[fixedImage,movingImage,weight,radius]". This
      metric is deprecated.  It still ``works'' but CC is preferred.

    \item Mean square difference: \verb"-m MSQ[fixedImage,movingImage,weight,0]". This metric works for intra-modality image registration. The last parameter \verb"0" is a padding value of no real meaning. For example, \verb"-m MSQ[fixed.nii,moving.nii,1,0]".

\end{itemize}

ANTS also support registration of point sets. The supported formats for point sets can be found in I/O section. The similarity metrics for point sets are:
\begin{itemize}
    \item Point set expectation: 
     \begin{verbatim}
-m PSE [fixedImage,movingImage,fixedPoints,movingPoints,weight,
pointSetPercentage,pointSetSigma,boundaryPointsOnly,
kNeighborhood,PartialMatchingIterations=100000]
    \end{verbatim}
    \begin{itemize}
        \item \verb"fixedImage": defines the space domain of the fixed point set.
        \item \verb"movingImage": defines the space domain of the moving point set.
        \item \verb"fixedPoints/Image": defines the coordinates of the fixed point set or label image. It can be an image with discrete positive labels, a VTK format point set file, or a text file. Details can be found in I/O section (TODO).
        \item \verb"movingPoints/Image": defines the coordinates of the moving point set or label image. 
        \item \verb"weight": weight for this metric. \verb"1" weights are relative to the weights on the $N$ other metrics passed to ANTS --- $N$ is unlimited.
        \item \verb"pointSetPercentage": the percentage of points to be randomly sampled used in the registration.
        \item \verb"pointSetSigma": the standard deviation of the Parzen window used to estimate the expectation. 
        \item \verb"boundaryPointsOnly":  \verb"1" (or ``true'') means only the boundary points in the label image is used to drive registration.
        \item \verb"kNeighborhood" is a positive discrete number. The first $k$ neighbors are used to compute the deformation during the registration. 
        \item \verb"PartialMatchingIterations" controls the symmtry in the matching. This option assumes the complete labeling is in the first set of label parameters ... more iterations leads to more symmetry in the matching  - 0 iterations means full asymmetry 
    \end{itemize}

    \item Jensen-Tsallis BSpline
    \begin{verbatim}
-m JTB[fixedImage,movingImage,fixedPoints,movingPoints,weight,
pointSetPercentage,pointSetSigma,boundaryPointsOnly,kNeighborhood,alpha,
meshResolution,splineOrder,numberOfLevels,useAnisotropicCovariances] 
    \end{verbatim}
    \begin{itemize}
        \item \verb"fixedImage": defines the space domain of the fixed point set.
        \item \verb"movingImage": defines the space domain of the moving point set.
        \item \verb"fixedPoints": defines the coordinates of the fixed point set. It can be an image with discrete positive labels, a VTK format point set file, or a text file. Details can be found in I/O section (TODO).
        \item \verb"movingPoints": defines the coordinates of the moving point set.
        \item \verb"weight": weight for this metric.  The weights are relative to the weights on the $N$ other metrics passed to ANTS --- $N$ is unlimited.
        \item \verb"pointSetPercentage": the percentage of points to be randomly sampled used in the registration.
        \item \verb"pointSetSigma": the sigma for the Parzen window used to estimate probabilities.  
        \item \verb"boundaryPointsOnly": [TODO] \verb"1" (or ``true'') means only the boundary points in the point sets are used to drive registration.
        \item \verb"kNeighborhood" is a positive discrete number. The first $k$ neighbors are used to compute the deformation during the registration. 
        \item \verb"alpha"
        \item \verb"meshResolution"
        \item \verb"splineOrder"
        \item \verb"numberOfLevels"
        \item \verb"useAnisotropicCovariances"
    \end{itemize}

\end{itemize}

ANTS has three affine registration similarity metrics for volumetric images, which are specified using \verb"--affine-metric-type":
\begin{itemize}
    \item Mutual information, specified as \verb"--affine-metric-type MI". This usually works for both inter and intra-modalities in 3D. Also, the options used in computing mututal information in affine registration can be specified as \verb"--MI-option N1xN2". The first parameter \verb"N1" specifies the number of bins. The second parameter \verb"N2" specifies the number of samples. For example: \verb"--MI-option 32x8000".
    \item Cross-correlation  \verb"--affine-metric-type CC".  CC has a
      wide capture range.  
    \item Mean square difference, specified using \verb"--affine-metric-type MSE". There is no options for this metric.
\end{itemize}

Fig. \ref{fig:metric_example} shows the registration result using intensity difference with the following command:
\begin{verbatim}
ANTS 2 -m MSQ[r16slice.nii,r64slice.nii,1,0] -r Gauss[3,0] -t SyN[0.5] -i 50x50x30
\end{verbatim}

%        \begin{wrapfigure}{r}{0.5\textwidth}
\begin{figure}
    \label{fig:metric_example}
    \centering
    \begin{tabular}[h]{c|c|c}
%        \includegraphics[width=0.16\textwidth]{Figures/r16slice.jpg} &
%        \includegraphics[width=0.16\textwidth]{Figures/r64slice.jpg} &
 %       \includegraphics[width=0.16\textwidth]{Figures/resMSQ.jpg} \\
        fixed image &
        moving image & 
        MSE  \\
    \end{tabular} 
    \itkcaption{registration using mean square intensity difference}
\end{figure}
    %    \end{wrapfigure}
Here is also an example script to register a pair of images using mean square intensity difference and computing the metrics of the registration image.
\begin{verbatim}
#use intensity difference with radius 0 -- radius no effect on intensity difference
ANTS 2 -m MSQ[r16slice.nii,r64slice.nii,1,0] -r Gauss[3,0] -t SyN[0.5] -i 50x50x30
WarpImageMultiTransform 2 r64slice.nii resMSQ.nii Warp.nii Affine.txt
MeasureImageSimilarity 2 0 r16slice.nii resMSQ.nii metricexamplelog.txt
MeasureImageSimilarity 2 1 r16slice.nii resMSQ.nii metricexamplelog.txt
MeasureImageSimilarity 2 2 r16slice.nii resMSQ.nii metricexamplelog.txt
ConvertToJpg resMSQ.nii resMSQ.jpg
\end{verbatim}

@@

\item Copy the file ANTS/Scripts/buildtemplateparallel.sh into that directory.  

@@

\verb  bash ~ \verb buildtemplateparallel.sh ~ \verb -h ~ to get usage. 

@@

The script offers an option ( -c 1  for qsub , -c 2 to attempt to use
multiple cores on your machine )  so that you can send the whole thing to distributed computing. 

@@

This succeeded for all subjects but one.  This subject had severe neurodegeneration which caused "under-normalization" to result.  A larger deformation mapping was required and so we increased the maximum number of iterations (the "-i" parameter).  One might also increase the gradient-descent step-size (SyN[0.25] $\rightarrow$ SyN[0.5]).   Large deformation mapping is challenging not only because of the amount of deformation, but also because the "true" solution is difficult to find.  Keeping this in mind, we also increased the "span" (radius) of the correlation window in the similarity metric (which also increases computation time). These modified parameters succeeded: 

@@

\verb Atropos ~tool, which is still under development.  We apply 

@@

      -c [10,0.001] -m [0.1,1x1x1] -i kmeans[3] -x mask.nii.gz

@@

The document atropos.pdf in the Documentation directory provides much
more explanation.

@@

\itkcaption[ANTS Segmentation]{The ANTs program {\em Atropos} extends basic Markov Random Field regularized Gaussian model tissue segmentation to include priors,

@@

  segmentation.  The ANTS program N4BiasFieldCorrection is 

@@

\item Preprocess images -- bias correction (e.g. via the \verb N4BiasFieldCorrection ~program), 
segmentation (via \verb Atropos ), etc.  Potentially construct an optimal template. 
